---
""
#author: Evgeny Galimov
""
date: '2022-11-24'
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(car)
library(jtools)
library(splines)
library(caret)
library(arm)
library(pROC)
library(predtools)
```

# loading data formatting variables and preprocessing 
```{r pressure, echo=FALSE}
final_table4 = read.csv('S://NDL/MouleshWorkings/Extracts2/PAM_carer_start_adjusted.csv')

final_table4 = final_table4[!is.na(final_table4$PAM_Value),]

final_table4 = add_column(final_table4,carer = ifelse(final_table4$MatchedPatientKey == 'carer',1, 0), .after = 'group' )

final_table4$PAM_Value = as.numeric(final_table4$PAM_Value)
final_table4$IMDDecile = as.numeric(  as.character(final_table4$IMDDecile )  )
final_table4$age = as.numeric(  as.character(final_table4$age )  )
final_table4$Gender = as.factor(final_table4$Gender)
final_table4$EthnicCategory = as.factor(final_table4$EthnicCategory)

# remove irrelevant data
final_table4 = final_table4[!final_table4$IMDDecile==-2,]
final_table4 = final_table4[!final_table4$Gender=='Other',]
```


## Path for loading
```{r pressure, echo=FALSE}
path = 'S://NDL/MouleshWorkings/2_efi_pam_bmi/pam/'
```

## Explore carer VS efi categories
```{r pressure, echo=FALSE}
summary(as.factor( final_table4$carer) )
summary( as.factor(final_table4$pam_cat_above50) )
```


## create binary outcome for PAM counts
```{r pressure, echo=FALSE}


final_table4$pam_cat_above60 = ifelse(final_table4$PAM_Value>60,1,0)
final_table4$pam_cat_above55 = ifelse(final_table4$PAM_Value>55,1,0)
final_table4$pam_cat_above50 = ifelse(final_table4$PAM_Value>50,1,0)
final_table4$pam_cat_above30 = ifelse(final_table4$PAM_Value>30,1,0)
```

# some descriptive graphs to compare group PAM split 50
```{r pressure, echo=FALSE}
if (!file.exists(path)) {dir.create(path)}




p1 = ggplot(final_table4,aes(x = PAM_Value)) + geom_histogram(binwidth=1) + xlab('PAM') #+  facet_wrap(~efi15)
png(paste(path,'0_PAM_hist.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p1)
dev.off()



p1 = ggplot(final_table4,aes(x = age)) + geom_histogram(binwidth=5) + xlab('Age') + facet_wrap(~pam_cat_above50)
png(paste(path,'0_age_splits50.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p1)
dev.off()

p2 = ggplot(final_table4,aes(x = Gender)) + geom_bar() + xlab('Gender') +  facet_wrap(~pam_cat_above50)
png(paste(path,'0_gender_splits50.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p2)
dev.off()

p3 = ggplot(final_table4, aes(x = IMDDecile)) +
  geom_bar() + xlab('IMDDecile') +
  scale_x_continuous(breaks = scales::pretty_breaks(n= 10)) +
  facet_wrap(~pam_cat_above50)+
  labs(x = 'IMDDecile', y = 'Counts') 

png(paste(path,'0_imd_splits50.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p3)
dev.off()

p4 = ggplot(final_table4, aes(x = EthnicCategory)) + geom_bar() + xlab('EthnicCategory') +
  facet_wrap(~pam_cat_above50)+ theme(axis.text.x=element_text(angle=15))
png(paste(path,'0_ethnicity_splits50.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p4)
dev.off()

```


# function to get linear model outputs
```{r pressure, echo=FALSE}
run_and_save_the_model = function(data, path, form_str, model_name, res_plots_f, shapirotest_f){

if (!file.exists(path)) {dir.create(path)}
m1 = lm(form_str, data = data)


sink(file = paste(path,model_name,'_summary.txt', sep = ''))
print( summ(m1, exp=F, confint=TRUE, digits = 5, ciwidth=0.05 ) )
sink()

png(paste(path,model_name,'_plot1_res.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,1)
dev.off()
png(paste(path,model_name,'_plot2_QQ.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,2)
dev.off()
png(paste(path,model_name,'_plot3_scale_locs.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,3)
dev.off()
png(paste(path,model_name,'_plot4_lev.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,5)
dev.off()

if (res_plots_f ==T) {
    crPlots(m1)
    resplot = recordPlot()
    png(paste(path,model_name,'_plot5_crPlots.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
    print(resplot)
    dev.off() }

if (shapirotest_f ==T) {
  
    high_lev_thresh = function(fit){ 
    p <- length(coefficients(fit))
    n <- length(fitted(fit))
    high_leverage_threshold = 2*(p+1)/n
    out = data.frame(high_leverage_threshold = c(high_leverage_threshold))
    return(out)}

    cooks_distance_thresh = function(fit){ 
    p <- length(coefficients(fit))
    n <- length(fitted(fit))
    cooks_distance_thresh_t = 4/(n-p-1)
    out = data.frame(cooks_distance_thresh = c(cooks_distance_thresh_t))
    return(out)}
    ## check normality of residuals
    # if p value > 0.05, distribution can be considered as normal
    stest = shapiro.test(m1$residuals)
    h_lev = high_lev_thresh(m1)
    cooks_d = cooks_distance_thresh(m1)
    
    df2 = setNames( data.frame(matrix(ncol = 3, nrow = 0)), c('c1','c2','c3')   )
    df2[nrow(df2) + 1,] = c( 'High leverage',h_lev, '')
    df2[nrow(df2) + 1,] = c( '','','')
    
    df2[nrow(df2) + 1,] = c( 'Cooks distance ',cooks_d, '')
    df2[nrow(df2) + 1,] = c( '','','')
    
    df2[nrow(df2) + 1,] = c( 'Shapiro test ',cooks_d, '')
    df2[nrow(df2) + 1,] = c( 'W',stest[1],'')
    df2[nrow(df2) + 1,] = c( 'p',stest[2],'')
    
    if (stest[2] < 0.05) {out = 'Residuals are not normally distributed'}
    if (stest[2] >= 0.05) {out = 'Residuals are normally distributed'}
    df2[nrow(df2) + 1,] = c( 'Conclusion:',out,'')
    
    write.csv(df2, paste(path,model_name,'_shapiro_cooks_leverage.csv', sep = '') )
    }


}
```


#  function to get model with splines outputs
```{r pressure, echo=FALSE}
run_and_save_the_model2 = function(data, path, form_str, model_name){

if (!file.exists(path)) {dir.create(path)}
m1 = lm(form_str, data = data)


sink(file = paste(path,model_name,'_summary_splines.txt',sep = ''))
print( summary(m1) )
 
sink()

png(paste(path,model_name,'_plot1_res.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,1)
dev.off()
png(paste(path,model_name,'_plot2_QQ.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,2)
dev.off()
png(paste(path,model_name,'_plot3_scale_locs.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,3)
dev.off()
png(paste(path,model_name,'_plot4_lev.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,5)
dev.off()


}
```

# run base linear model: PAM_Value ~ carer+Gender+age+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}
res_plots_f = T
shapirotest_f = T 
model_name = '1_PAM_base_lm'
form_str = as.formula('PAM_Value ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model(final_table4, path, form_str, model_name, T, T)
```

# run linear model where Y is modelled as log: log(PAM_Value_non0) ~ carer+Gender+age+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}
res_plots_f = T
shapirotest_f = T 
# convert PAM score 0 to 0.001
final_table4$PAM_Value_non0 = ifelse(final_table4$PAM_Value==0, 0.001, final_table4$PAM_Value)
model_name = '2_pam_base_lm_logY'
form_str = as.formula('log(PAM_Value_non0) ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model(final_table4, path, form_str, model_name, T, T)
```

# run linear model where Y is modelled as sqrt: sqrt(PAM_Value) ~ carer+Gender+age+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}
res_plots_f = T
shapirotest_f = T
model_name = '3_pam_base_lm_sqrtY'
form_str = as.formula('sqrt(PAM_Value) ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model(final_table4, path, form_str, model_name,T,T)
```

# run linear model where age is exponentiated: PAM_Value ~ carer+Gender+exp(0.03*age)+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}
res_plots_f = F
shapirotest_f = F 
model_name = '4_pam_base_lm_expAge'
form_str = as.formula('PAM_Value ~ carer+Gender+exp(0.03*age)+IMDDecile+EthnicCategory')
run_and_save_the_model(final_table4, path, form_str, model_name,T,T)
```

# run linear model where where Y is modelled as log and age is exponentiated: log(PAM_Value) ~ carer+Gender+exp(0.03*age)+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}
res_plots_f = F
shapirotest_f = F 
model_name = '5_pam_base_lm_expAge_logY'
form_str = as.formula('log(PAM_Value) ~ carer+Gender+exp(0.03*age)+IMDDecile+EthnicCategory')
run_and_save_the_model(final_table4, path, form_str, model_name,T,T)
```

# run linear model where where Y is modelled as log and with splines for age: log(PAM_Value) ~ carer+Gender+exp(0.03*age)+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}

model_name = '9_pam_lm_splines3_age'
form_str = as.formula('PAM_Value ~ carer+Gender+bs(age)+IMDDecile+EthnicCategory')
run_and_save_the_model2(final_table4, path, form_str, model_name)
```



# function for running logistic regression
```{r pressure, echo=FALSE}
run_and_save_the_model_logr = function(data, out_true, path, form_str, model_name){

if (!file.exists(path)) {dir.create(path)}
m1 = glm(form_str, data = data, family=binomial)
sink(file = paste(path,model_name,'_summary.txt', sep = ''))
print( summ(m1, exp=T, digits = 5, confint=TRUE, ciwidth=0.05 ) )
sink()


# conf matrix
  pred = as.factor(ifelse(fitted(m1) >=.5, 1, 0))
  if ( (length(levels(pred))==1) & (levels(pred)[1]==1) ) {levels(pred)[length(levels(pred))+1] = '0'}
  if ( (length(levels(pred))==1) & (levels(pred)[1]==0) ){levels(pred)[length(levels(pred))+1] = '1'}
  

sink(file = paste(path,model_name,'_conf.matrix.txt', sep = ''))
print(   confusionMatrix(pred, as.factor(out_true) )   )
sink()

binnedplot(fitted(m1), residuals(m1, type = 'response'), 
           nclass = NULL,xlab='Expected values',ylab='Average residual',
           main='Binned residual plot',
           cex.pts=0.8,col.pts=1,col.int='grey')
b1 = recordPlot()
png(paste(path,model_name,'_plot1_binned_plot.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
    print(b1)
    dev.off()

# roc auc
      # out_true = efi$efi15 (0,1 - numeric)
invisible(  plot(    roc(out_true, fitted(m1)   ),print.auc=T, col='red'      )  )
pauc = recordPlot()
png(paste(path,model_name,'_plot2_roc_auc.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
    print(pauc)
    dev.off()

    
# calibration plot
data_plot = data

data_plot$y = out_true
data_plot$pred = predict.glm(m1, type = 'response')


c_plot = calibration_plot(data_plot, obs='y',  pred='pred' )
#c_plot = recordPlot()
png(paste(path,model_name,'_plot3_calibration_plot.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
print(c_plot)
dev.off()  
    
}
```






# run logistic regression for variious threshold for splitting PAM score
```{r pressure, echo=FALSE}
form_str = as.formula('pam_cat_above60 ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(final_table4, final_table4$pam_cat_above60, path, form_str, '5_logr_pam_above60')
form_str = as.formula('pam_cat_above55 ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(final_table4, final_table4$pam_cat_above55, path, form_str, '6_logr_pam_above55')
form_str = as.formula('pam_cat_above50 ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(final_table4, final_table4$pam_cat_above50, path, form_str, '6_logr_pam_above50')
form_str = as.formula('pam_cat_above30 ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(final_table4, final_table4$pam_cat_above30, path, form_str, '7_logr_pam_above30')

```
























