---
title: '1'
output: html_document
date: '2022-11-24'
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(car)
library(jtools)
library(splines)
library(caret)
library(arm)
library(pROC)
library(predtools)

```

```{r pressure, echo=FALSE}
par(mar = c(1,1,1,1))

```


## Path for loading
```{r pressure, echo=FALSE}
path = 'S://NDL/MouleshWorkings/Extracts2/'
```


# load data
```{r pressure, echo=FALSE}
efi = read.csv('S://NDL/MouleshWorkings/230110_EFI_adjusted.csv')

efi$carer = ifelse(efi$Carer_Flag=='Carer',1,0)
efi$carer = as.factor(efi$carer)
efi$Gender = as.factor(efi$Gender)
efi$age = as.numeric(  as.character(efi$age )  )
efi$IMDDecile = as.numeric(  as.character(efi$IMDDecile )  )
efi$EthnicCategory = as.factor(efi$EthnicCategory)

# remove irrelevant data
efi = efi[!efi$IMDDecile==-2,]
efi = efi[!efi$Gender=='Other',]
efi = efi[!efi$Gender=='Unknown',]
efi$Gender = droplevels(efi$Gender)
```


```{r pressure, echo=FALSE}

summary(efi$carer) 

summary(efi$Gender) 

summary(efi$IMDDecile) 

summary(efi$age) 

summary(efi$EthnicCategory) 

summary(efi$Count)

```




## set path for saving and save the data
```{r pressure, echo=FALSE}
path = 'S://NDL/MouleshWorkings/2_efi_pam_bmi/efi/'
write.csv(efi, paste(path, 'data.csv', sep=''))

```

## create subsets of efi by counts
```{r pressure, echo=FALSE}
efi_15 = efi[efi$efiCountNumber<16,]
efi_16 = efi[efi$efiCountNumber<17,]
efi_17 = efi[efi$efiCountNumber<18,]
efi_18 = efi[efi$efiCountNumber<18,]
efi_19 = efi[efi$efiCountNumber<18,]

efi_15a = efi[efi$efiCountNumber>=16,]
efi_16a = efi[efi$efiCountNumber>=17,]
efi_17a = efi[efi$efiCountNumber>=18,]
efi_18a = efi[efi$efiCountNumber>=18,]
efi_19a = efi[efi$efiCountNumber>=18,]

```


## create binary outcome for efi counts
```{r pressure, echo=FALSE}
# print( paste('Fit: ', min(efi[efi$ElectronicFrailtyIndex=='Fit',]$efiCountNumber), '-', max(efi[efi$eFICategory=='Fit',]$efiCountNumber))  )
# print( paste('Mild: ', min(efi[efi$ElectronicFrailtyIndex=='Mild',]$efiCountNumber), '-', max(efi[efi$eFICategory=='Mild',]$efiCountNumber))  )
# print( paste('Moderate: ', min(efi[efi$ElectronicFrailtyIndex=='Moderate',]$efiCountNumber), '-', max(efi[efi$eFICategory=='Moderate',]$efiCountNumber))  )
# print( paste('Severe: ', min(efi[efi$ElectronicFrailtyIndex=='Severe',]$efiCountNumber), '-', max(efi[efi$eFICategory=='Severe',]$efiCountNumber))  )

efi$efi_cat_mild = ifelse(efi$Count>4,1,0)
efi$efi_cat_moderate = ifelse(efi$Count>7,1,0)
efi$efi_cat_severe = ifelse(efi$Count>11,1,0)

efi$efi15 = ifelse(efi$Count>15,1,0)
efi$efi16 = ifelse(efi$Count>16,1,0)
efi$efi17 = ifelse(efi$Count>17,1,0)
efi$efi18 = ifelse(efi$Count>18,1,0)
efi$efi19 = ifelse(efi$Count>19,1,0)

#efi_15 = efi[efi$efiCountNumber<16,]
#efi_16 = efi[efi$efiCountNumber<17,]
#efi_17 = efi[efi$efiCountNumber<18,]
#efi_18 = efi[efi$efiCountNumber<18,]
#efi_19 = efi[efi$efiCountNumber<18,]






```

## Explore carer VS efi categories
```{r pressure, echo=FALSE}
summary(efi$carer)
summary(efi$efi_cat_moderate)

table(efi$carer, efi$efi_cat_moderate)
prop.table( table(efi$carer, efi$efi_cat_moderate) )


```



# some descriptive graphs to compare group with eFI <=15 and more than 15
```{r pressure, echo=FALSE}
p1 = ggplot(efi,aes(x = Count)) + geom_histogram(binwidth=1) + xlab('eFI')
png(paste(path,'0_efi_counts.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p1)
dev.off()


p1 = ggplot(efi,aes(x = Count)) + geom_histogram(binwidth=1) + xlab('eFI') +
  facet_wrap(~efi_cat_moderate)
png(paste(path,'0_efi_splits_mod.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p1)
dev.off()



p1 = ggplot(efi,aes(x = age)) + geom_histogram(binwidth=5) + xlab('Age') +
  facet_wrap(~efi_cat_moderate)
png(paste(path,'0_age_splits_mod.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p1)
dev.off()

p2 = ggplot(efi,aes(x = Gender)) + geom_bar() + xlab('Gender') +  facet_wrap(~efi_cat_moderate)
png(paste(path,'0_gender_splits_mod.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p2)
dev.off()

p3 = ggplot(efi, aes(x = IMDDecile)) +
  geom_bar() + xlab('IMDDecile') +
  scale_x_continuous(breaks = scales::pretty_breaks(n= 10)) +
  facet_wrap(~efi_cat_moderate)+
  labs(x = 'IMDDecile', y = 'Counts') 
  #theme(axis.text = element_text(size = 20))+
  #theme(axis.title = element_text(size = 24, face = 'bold'))
png(paste(path,'0_imd_splits_mod.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p3)
dev.off()

p4 = ggplot(efi,aes(x = EthnicCategory)) + geom_bar() + xlab('EthnicCategory') +
  facet_wrap(~efi_cat_moderate)+ theme(axis.text.x=element_text(angle=15))
png(paste(path,'0_ethnicity_splits_mod.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
print(p4)
dev.off()

```


# function to get linear model outputs
```{r pressure, echo=FALSE}
run_and_save_the_model = function(data, path, form_str, model_name, res_plots_f, shapirotest_f){

if (!file.exists(path)) {dir.create(path)}
m1 = lm(form_str, data = data)


sink(file = paste(path,model_name,'_summary.txt', sep = ''))
print( summ(m1, exp=F, confint=TRUE, digits = 5, ciwidth=0.05 ) )
sink()

png(paste(path,model_name,'_plot1_res.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,1)
dev.off()
png(paste(path,model_name,'_plot2_QQ.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,2)
dev.off()
png(paste(path,model_name,'_plot3_scale_locs.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,3)
dev.off()
png(paste(path,model_name,'_plot4_lev.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,5)
dev.off()

if (res_plots_f ==T) {
    crPlots(m1)
    resplot = recordPlot()
    png(paste(path,model_name,'_plot5_crPlots.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
    print(resplot)
    dev.off() }

if (shapirotest_f ==T) {
  
    high_lev_thresh = function(fit){ 
    p <- length(coefficients(fit))
    n <- length(fitted(fit))
    high_leverage_threshold = 2*(p+1)/n
    out = data.frame(high_leverage_threshold = c(high_leverage_threshold))
    return(out)}

    cooks_distance_thresh = function(fit){ 
    p <- length(coefficients(fit))
    n <- length(fitted(fit))
    cooks_distance_thresh_t = 4/(n-p-1)
    out = data.frame(cooks_distance_thresh = c(cooks_distance_thresh_t))
    return(out)}
    
    ## check normality of residuals
    # if p value > 0.05, distribution can be considered as normal
    stest = shapiro.test(m1$residuals)
    h_lev = high_lev_thresh(m1)
    cooks_d = cooks_distance_thresh(m1)
    
    df2 = setNames( data.frame(matrix(ncol = 3, nrow = 0)), c('c1','c2','c3')   )
    df2[nrow(df2) + 1,] = c( 'High leverage',h_lev, '')
    df2[nrow(df2) + 1,] = c( '','','')
    
    df2[nrow(df2) + 1,] = c( 'Cooks distance ',cooks_d, '')
    df2[nrow(df2) + 1,] = c( '','','')
    
    df2[nrow(df2) + 1,] = c( 'Shapiro test ',cooks_d, '')
    df2[nrow(df2) + 1,] = c( 'W',stest[1],'')
    df2[nrow(df2) + 1,] = c( 'p',stest[2],'')
    
    if (stest[2] < 0.05) {out = 'Residuals are not normally distributed'}
    if (stest[2] >= 0.05) {out = 'Residuals are normally distributed'}
    df2[nrow(df2) + 1,] = c( 'Conclusion:',out,'')
    
    write.csv(df2, paste(path,model_name,'_shapiro_cooks_leverage.csv', sep = '') )
    }
}
```


#  function to get model with splines outputs
```{r pressure, echo=FALSE}
run_and_save_the_model2 = function(data, path, form_str, model_name){

if (!file.exists(path)) {dir.create(path)}
m1 = lm(form_str, data = data)


sink(file = paste(path,model_name,'_summary_splines.txt',sep = ''))
print( summary(m1) )
 
sink()

png(paste(path,model_name,'_plot1_res.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,1)
dev.off()
png(paste(path,model_name,'_plot2_QQ.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,2)
dev.off()
png(paste(path,model_name,'_plot3_scale_locs.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,3)
dev.off()
png(paste(path,model_name,'_plot4_lev.png', sep = ''), 
    width = 20, height = 15, units = 'cm', res = 200  )
plot(m1,5)
dev.off()


}
```

# run base linear model: Count ~ carer+Gender+age+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}
res_plots_f = F
shapirotest_f = F 
model_name = '1_efi_base_lm'
form_str = as.formula('Count ~ carer+Gender+age+IMDDecile')
run_and_save_the_model(efi, path, form_str, model_name,T,F)
```

# run linear model with splines for age: Count ~ carer+Gender+bs(age)+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}

model_name = '14_efi_lm_splines3_age'
form_str = as.formula('Count ~ carer+Gender+bs(age)+IMDDecile+EthnicCategory')
run_and_save_the_model2(efi, path, form_str, model_name)
```

# run linear model where Y is modelled as log: log(Count) ~ carer+Gender+age+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}
res_plots_f = F
shapirotest_f = F 
model_name = '2_efi_base_lm_logY'
form_str = as.formula('log(Count) ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model(efi, path, form_str, model_name,F,F)
```

# run linear model where Y is modelled as sqrt: sqrt(Count) ~ carer+Gender+age+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}
res_plots_f = F
shapirotest_f = F 
model_name = '3_efi_base_lm_sqrtY'
form_str = as.formula('sqrt(Count) ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model(efi, path, form_str, model_name,F,F)
```

# run linear model where Y is modelled as sqrt and age is exponentiated: sqrt(Count) ~ carer+Gender+age+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}
res_plots_f = F
shapirotest_f = F 
model_name = '4_efi_base_lm_expAge'
form_str = as.formula('sqrt(Count) ~ carer+Gender+exp(0.03*age)+IMDDecile+EthnicCategory')
run_and_save_the_model(efi, path, form_str, model_name,F,F)
```

# run linear model where Y is modelled as log and age is exponentiated: log(Count) ~ carer+Gender+exp(0.03*age)+IMDDecile+EthnicCategory
```{r pressure, echo=FALSE}
res_plots_f = F
shapirotest_f = F 
model_name = '5_efi_base_lm_expAge_logY'
form_str = as.formula('log(Count) ~ carer+Gender+exp(0.03*age)+IMDDecile+EthnicCategory')
run_and_save_the_model(efi, path, form_str, model_name,F,F)
```


# function for running logistic regression
```{r pressure, echo=FALSE}
run_and_save_the_model_logr = function(data, out_true, path, form_str, model_name){

if (!file.exists(path)) {dir.create(path)}
m1 = glm(form_str, data = data, family=binomial)

sink(file = paste(path,model_name,'_summary.txt', sep = ''))
print( summ(m1, exp=T, digits = 5, confint=TRUE, ciwidth=0.05 ) )
 
sink()


# conf matrix
  pred = as.factor(ifelse(fitted(m1) >=.5, 1, 0))
  if ( (length(levels(pred))==1) & (levels(pred)[1]==1) ) {levels(pred)[length(levels(pred))+1] = '0'}
  if ( (length(levels(pred))==1) & (levels(pred)[1]==0) ){levels(pred)[length(levels(pred))+1] = '1'}
  

sink(file = paste(path,model_name,'_conf.matrix.txt', sep = ''))
print(   confusionMatrix(pred, as.factor(out_true) )   )
sink()

binnedplot(fitted(m1), residuals(m1, type = 'response'), 
           nclass = NULL,xlab='Expected values',ylab='Average residual',
           main='Binned residual plot',
           cex.pts=0.8,col.pts=1,col.int='grey')
b1 = recordPlot()
png(paste(path,model_name,'_plot1_binned_plot.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
    print(b1)
    dev.off()

# roc auc
      # out_true = efi$efi15 (0,1 - numeric)
invisible(  plot(    roc(out_true, fitted(m1)   ),print.auc=T, col='red'      )  )
pauc = recordPlot()
png(paste(path,model_name,'_plot2_roc_auc.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
    print(pauc)
    dev.off()

# calibration plot
data_plot = data
#data_plot$pred = predict.glm(m1, type='response')
#calibration_plot(data = data_plot, obs=out_true, pred=pred,'Calibration plot')

data_plot$y = out_true
data_plot$pred = predict.glm(m1, type = 'response')


c_plot = calibration_plot(data_plot, obs='y',  pred='pred' )
#c_plot = recordPlot()
png(paste(path,model_name,'_plot3_calibration_plot.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
print(c_plot)
dev.off()    
    
}
```



# run logistic regression for variious threshold for splitting EFI
```{r pressure, echo=FALSE}
form_str = as.formula('efi15 ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(efi, efi$efi15, path, form_str, '6_logr_efi15')
form_str = as.formula('efi16 ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(efi, efi$efi16, path, form_str, '7_logr_efi16')
form_str = as.formula('efi17 ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(efi, efi$efi17, path, form_str, '8_logr_efi17')
form_str = as.formula('efi18 ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(efi, efi$efi18, path, form_str, '9_logr_efi18')
form_str = as.formula('efi19 ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(efi, efi$efi19, path, form_str, '10_logr_efi19')
form_str = as.formula('efi_cat_mild ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(efi, efi$efi_cat_mild, path, form_str, '11_logr_efi_mild')
form_str = as.forsula('efi_cat_moderate ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(efi, efi$efi_cat_moderate, path, form_str, '12_logr_efi_moderate')
form_str = as.formula('efi_cat_severe ~ carer+Gender+age+IMDDecile+EthnicCategory')
run_and_save_the_model_logr(efi, efi$efi_cat_severe, path, form_str, '13_logr_efi_severe')

```




# save figures for '12_logr_efi_moderate' log model

```{r pressure, echo=FALSE}
m1 = glm(efi_cat_moderate ~ carer+Gender+age+IMDDecile+EthnicCategory, data = efi, family=binomial)

model_name = '12_logr_efi_moderate'

sink(file = paste(path,model_name,'_summary.txt', sep = ''))
print( summ(m1, exp=T, digits = 5, confint=TRUE, ciwidth=0.05 ) )
sink()
# conf matrix
pred = as.factor(ifelse(fitted(m1) >=.5, 1, 0))
if ( (length(levels(pred))==1) & (levels(pred)[1]==1) ) {levels(pred)[length(levels(pred))+1] = '0'}
if ( (length(levels(pred))==1) & (levels(pred)[1]==0) ){levels(pred)[length(levels(pred))+1] = '1'}
```

###Binned plot
```{r pressure, echo=FALSE}
path = 'S://NDL/MouleshWorkings/2_efi_pam_bmi/efi0/'

model_name = '12_logr_efi_moderate'

x = fitted(m1)
n <- length(x)
nclass = NULL
if (is.null(nclass)) {
  if (n >= 100) {
    nclass = floor(sqrt(length(x)))
  }
  if (n > 10 & n < 100) {
    nclass = 10
  }
  if (n <= 10) {
    nclass = floor(n/2)
  }
}

aa = data.frame(binned.resids(fitted(m1), residuals(m1, type = 'response'), nclass)$binned)
write.csv(aa, paste(path,model_name,'_binned_plot.csv', sep = ''))
xlab='Expected values'
ylab='Average residual'
main='Binned residual plot'
cex.pts=0.8
col.pts=1
col.int='grey'

plot(range(aa$xbar), range(aa$ybar, aa$X2se, -aa$X2se, na.rm = TRUE), 
     xlab = xlab, ylab = ylab, type = "n", cex.lab=1.5,cex.axis=1.2)
abline(0, 0, lty = 2)
lines(aa$xbar, aa$X2se, col = col.int)
lines(aa$xbar, -aa$X2se, col = col.int)
points(aa$xbar, aa$ybar, pch = 19, cex = cex.pts, col = col.pts)
p1 = recordPlot()
png(paste(path,model_name,'_binned_plot.png', sep = ''), 
        width = 15, height = 10, units = 'cm', res = 300  )
print(p1)
dev.off() 

```


###AUC plot
```{r pressure, echo=FALSE}

out_true = efi$efi_cat_moderate
invisible(  plot(    roc(out_true, fitted(m1)   ),print.auc=T, col='red',cex.lab=1.5,cex.axis=1.2    )  )
pauc = recordPlot()
jpeg(paste(path,model_name,'_plot2_roc_auc.png', sep = ''), 
        width = 12, height = 10, units = 'cm', res = 300  )
    print(pauc)
    dev.off()
    
a = roc(out_true, fitted(m1))
df = as.data.frame(a$thresholds)
df$sensitivities = a$sensitivities 
df$specificities = a$specificities  
write.csv(df, paste(path,model_name,'_auc_plot.csv', sep = ''))


```








# Figures for final log model: 
```{r pressure, echo=FALSE}
m1 = glm(form_str, data = efi, family=binomial)

sink(file = paste(path,model_name,'_summary.txt', sep = ''))
print( summ(m1, exp=T, digits = 5, confint=TRUE, ciwidth=0.05 ) )
 
sink()


# conf matrix
  pred = as.factor(ifelse(fitted(m1) >=.5, 1, 0))
  if ( (length(levels(pred))==1) & (levels(pred)[1]==1) ) {levels(pred)[length(levels(pred))+1] = '0'}
  if ( (length(levels(pred))==1) & (levels(pred)[1]==0) ){levels(pred)[length(levels(pred))+1] = '1'}
  

sink(file = paste(path,model_name,'_conf.matrix.txt', sep = ''))
print(   confusionMatrix(pred, as.factor(out_true) )   )
sink()

binnedplot(fitted(m1), residuals(m1, type = 'response'), 
           nclass = NULL,xlab='Expected values',ylab='Average residual',
           main='Binned residual plot',
           cex.pts=0.8,col.pts=1,col.int='grey')
b1 = recordPlot()
png(paste(path,model_name,'_plot1_binned_plot.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
    print(b1)
    dev.off()

# roc auc
      # out_true = efi$efi15 (0,1 - numeric)
invisible(  plot(    roc(out_true, fitted(m1)   ),print.auc=T, col='red'      )  )
pauc = recordPlot()
png(paste(path,model_name,'_plot2_roc_auc.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
    print(pauc)
    dev.off()

# calibration plot
data_plot = efi
#data_plot$pred = predict.glm(m1, type='response')
#calibration_plot(data = data_plot, obs=out_true, pred=pred,'Calibration plot')

data_plot$y = out_true
data_plot$pred = predict.glm(m1, type = 'response')


c_plot = calibration_plot(data_plot, obs='y',  pred='pred' )
#c_plot = recordPlot()
png(paste(path,model_name,'_plot3_calibration_plot.png', sep = ''), 
        width = 20, height = 15, units = 'cm', res = 200  )
print(c_plot)
dev.off()    
```













